---
editor_options: 
  markdown: 
    wrap: 72
---

# 7.4 Introduction

# 7.4.1 Analysing Spatial Autocorrelation with Moran's I, LISA and friends

# 7.4.2 Data download

```{r}
library(here)
library(janitor)
library(sf)
library(tidyverse)
```

```{r}
# read the ward data in
LondonWards <- st_read(here::here("week7",
                                  "week07data",
                                  "London-wards-2018","London-wards-2018_ESRI","London_Ward.shp"
                                  ))
```

```{r}
LondonWardsMerged <- st_read(here::here("week7",
                                  "week07data",
                                  "London-wards-2018",
                                  "London-wards-2018_ESRI",
                                  "London_Ward_CityMerged.shp"))%>%
  st_transform(.,27700)
```

```{r}
wardData <- read_csv(here::here("week7",
                                  "week07data",
                                "ward-profiles-excel-version.csv"), 
                     locale = locale(encoding = 'latin1'),
                     na=c("NA","n/a"))%>%
  clean_names()
```

```{r}
LondonWardsMerged <- LondonWardsMerged %>%
  left_join(wardData,
            by=c("GSS_CODE"="new_code"))%>%
  dplyr::distinct(GSS_CODE,.keep_all=T)%>%
  dplyr::select(GSS_CODE, ward_name,average_gcse_capped_point_scores_2014)
# dplyr::distinct(GSS_CODE, .keep_all = T): This line uses the distinct function to keep only unique rows based on the "GSS_CODE" column. 
# The .keep_all = T argument ensures that all columns are retained for the distinct rows.
# dplyr::select() : selects specific columns 
```

```{r}
# Checking the projection system
st_crs(LondonWardsMerged)
```

```{r}
library(tmap)

BluePlaques <- st_read(here::here("Week6","week06data", 
                                  "open-plaques-london-2018-04-08.geojson"
                                  )) %>%
  st_transform(.,27700)
```

```{r}
tmap_mode("plot")
tm_shape(LondonWardsMerged)+
  tm_polygons(col = NA, alpha = 0.5)+
  tm_shape(BluePlaques)+
  tm_dots(col="blue")
```
# 7.5 Data cleaning
```{r}
summary(BluePlaques)
```

```{r}
# some points are located in outside of london, so need to cleaning the data
BluePlaquesSub <- BluePlaques[LondonWardsMerged,]
tm_shape(LondonWardsMerged)+
  tm_polygons(col=NA,alpha = 0.5)+
  tm_shape(BluePlaquesSub)+
  tm_dots(col="blue")
```
# 7.6 Data manipulation
## exmaple1: st_inetrsects(polygon,point)
```{r}
# To create a continuous observation from the blue plaques data
# to count all of the blue plaques that fall within each Ward in the City.
# Using st_intersects() function from the sf package

example <- st_intersects(LondonWardsMerged,BluePlaquesSub)

example
```
## exmaple2: st_join() and filter
```{r}
check_example <- LondonWardsMerged %>%
  st_join(BluePlaquesSub)%>%
# Performs a spatial join using the st_join function. It joins the LondonWardsMerged data frame with the BluePlaquesSub spatial dataset based on their spatial relationships.
  filter(ward_name=="Kingston upon Thames - Coombe Hill")
# Choosing the data of Kingston upon Thames - Coombe Hill
```

## calculating the number of intersection in each ward

```{r}
library(sf)
point_sf_joined <- LondonWardsMerged%>%
  mutate(n=lengths(st_intersects(.,BluePlaquesSub)))%>%
  janitor::clean_names()

# 'lengths' This variable represents the number of intersections between each geometry in LondonWardsMerged and the geometries in the BluePlaquesSub dataset. 
# lengths() is a function in R that is used to get the lengths of elements in a list-like object. #  lengths() is then used to get the number of intersections for each row.

# janitor::clean_names() to lowercase/ replace spaces with underscores...
```

## calculating area (面积) and density and select the variables

```{r}

# calculate area 
# The st_area() function is used to calculate the area of each geometry in the spatial data frame. 
points_sf_joined <- point_sf_joined %>%
  mutate(area=st_area(.)) %>%
# then density of the points per ward
  mutate(density=n/area) %>%
# select density andd some other variables
  dplyr::select(density, ward_name, gss_code, n, average_gcse_capped_point_scores_2014)
```

## ploting and checking the data

```{r}

points_sf_joined_2<- points_sf_joined %>%                    
  group_by(gss_code) %>%         
  summarise(density = first(density),
          wardname= first(ward_name),
          plaquecount= first(n))
# # density = first(density): This extracts the first value of the density variable within each group.
```

```{r}
#example of using first() function
# Create a sample data frame
#df <- data.frame(
  #group = c("A", "A", "B", "B", "C", "C"),
  #value = c(10, 20, 15, 25, 30, 40)

# Use dplyr to group by 'group' and extract the first value of 'value' in each group
#result <- df %>%
  #group_by(group) %>%
  #summarise(first_value = first(value))

# Print the result
#print(result)

# results:
# A tibble: 3 × 2
#  group first_value
#  <chr>       <dbl>
#1 A              10
#2 B              15
#3 C              30
```

```{r}
#mapping
tm_shape(points_sf_joined_2) +
  tm_polygons("density",
              style="jenks",
              palette="PuOr",
              midpoint=NA,
              popup.vars=c("wardname","density"),
              title="Blue Plaque Density")
# popup.vars=c("wardname","density"): This specifies the variables to be shown in the pop-up when you click on the map. In this case, it would show the "wardname" and "density." 
```

# 7.7 Weight matrix

```{r}
# to calculate Moran'I and any similar statistics 
# need to define a W spatial weights matrix
library(spdep)
```

```{r}
#First calculate the centroids of all Wards in London
coordsW <- points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
# st_centroid(): This function calculates the centroid of the geometries in your sf object. The centroid is a point representing the center of mass of the geometry.
# st_geometry(): This extracts the geometry (in this case, the centroid) from the sf object.
plot(coordsW,axes=TRUE)
# The axes=TRUE argument specifies that axes should be drawn on the plot.
```
```{r}
# to generate a spatial weights matrix

# defined by neighboring places(lines,points,both) Usually defined: rook case (move fowards or sidewards), bishops case, queen's case

# Queen's case definition: polygons with a shared edge or a corner will be included in computations for the target polygon

# Alternatively instead of using contiguous relationships you can use distance based relationships.
```

```{r}
# reate a neighbours list — which is a list of all the neighbours.
LWard_nb <- points_sf_joined %>%
  poly2nb(.,queen = T)
# using poly2nb() with the argument queen=T saying we want a to use Queens case. 
```

```{r}
# summary
summary(LWard_nb)
# the average number of neighbours is 5.88
```

```{r}
# plot them 
plot(LWard_nb,st_geometry(coordsW),col="red")
# plotting the centroids (coordsW) on top of the map represented by LWard_nb. 

#add a map underneath (the base boundary map)
plot(points_sf_joined$geometry, add =T)
# The add=T argument indicates that you want to add this layer to the existing plot.
```
## 7.7.1 Matrix style
```{r}
# create a spatial weight matrix. 

# makes a matrix the size of the number of
# neighbourhoods The matrix style: 
# B is the basic binary coding (1/0) 
# W is row standardised (sums over all links to n) 
# C is globally standardised(sums over all links to n) 
# U is equal to C divided by the number of neighbours (sums over all links to unity) 
# S is the variance-stabilizing coding scheme proposed by Tiefelsdorf et al. 1999, p. 167-168 (sums over all links to n).
```

```{r}
# create a spatial weights matrix from these weights
Lward.lw <- LWard_nb %>%
  nb2mat(.,style="B") # binary (whether or not)
# using the nb2mat function to convert a spatial weights matrix to a matrix. 

sum(Lward.lw)
# The sum(Lward.lw) line is essentially counting the number of non-zero elements in the spatial weights matrix, indicating the number of neighbors each spatial unit has based on the specified spatial relationships.指示每个空间单元基于指定的空间关系具有的邻居数量。
```
```{r}
# In globally standardised (C), all weights are scaled so that the sum of all weights in the matrix is equal to n, the number of observations.
Lward.lw_C <- LWard_nb %>%
  nb2mat(.,style="C")
sum(Lward.lw_C)
```
```{r}
# each spatial weight has a value of x(0.1697051):
x=633/3730
print(x)
```
```{r}
# do row standardisation(W)
# the weights for each observation are scaled so that the sum of weights for each observation is equal to 1.
# 1 is divided by the sum of the number of neighbours in each row. 
sum(Lward.lw[2,])
# row 2 here sums to 3, meaning each weight would be 0.3333 in row 2 only.
```
```{r}
# Row standardization
# LWard_nb is spatial weights matrix
Lward.lw_W <- nb2mat(LWard_nb, style="W")
# This calculates the sum of weights for each row (observation) in the original matrix.
row_sums <- rowSums(Lward.lw_W)
# The rowSums() function in R is used to calculate the sum of values across rows in a matrix or a data frame.
# Perform row standardization
Lward.lw_W_row_standardized <- Lward.lw_W / row_sums
```
# 7.8 Autocorrelation
```{r}
# calculate the Moran’s I 
# Moran’s I requires a spatial weight list type object as opposed to matrix, 
Lward.list <- LWard_nb %>%
    nb2listw(., style="C")
# The nb2listw() function can be used to convert a spatial weights matrix to a list format. 
# style="C", which indicates global standardization.
```
## 7.8.1 Moran’s I
```{r}
# have clustered values (close to 1) or dispersed values (close to -1) 
# will calculate for the densities
I_LWard_Global_Density <- points_sf_joined_2 %>%
  # points_sf_joined: The data frame or a spatial object containing the variable of ("density").
  pull(density)%>%
  #The pull() function from the dplyr package is used to extract the column named "density" from your data.
  as.vector()%>%
  # as.vector(): Converts the extracted column to a vector.
  moran.test(.,Lward.list)
  # Performs the Moran's I test for spatial autocorrelation. 
  # moran.test(the vector of variable,the spatial weights list)
I_LWard_Global_Density

#results:
# p-value <0.05 reject the null hypothesis of no spatial autocorrelation. 
# Moran's I is 0.21317 positive spatial autocorrelation
# similar values are clustered in space. In other words, neighboring locations tend to have similar values of the "density" variable.
```
## 7.8.2 Geary’s C
```{r}
GC_LWard_Global_Density <- 
  points_sf_joined_2 %>%
  pull(density) %>%
  as.vector()%>%
  geary.test(.,Lward.list)
  
GC_LWard_Global_Density

# results:
# 0 < Geary C statistic  (0.78632658) <1 positive spatial autocorrelation 
#  (remember Geary’s C falls between 0 and 2???）
```
## 7.8.3 Getis Ord
```{r}
G_LWard_Global_Density <- 
  points_sf_joined_2 %>%
  pull(density)%>%
  as.vector()%>%
  globalG.test(.,Lward.list)

G_LWard_Global_Density

# results:
# Scientific notation
# Global G statistic(0.004732322) > Expectation(0.001582278)
# High values clustering;（hot spot)
```
## 7.8.4 Summary
## 7.8.5 Local Moran’s I
```{r}
# Local Moran’s I is:
# The difference between a value and neighbours * the sum of differences between neighbours and the mean
# Where the the difference between a value and neighbours is divided by the standard deviation (how much values in neighbourhood vary about the mean)

# Z score (standard deviation)
# how many standard deviations a value is away (above or below) from the mean. 
# z>1.65时呈现聚集分布，
# z<-1.65时呈现离散分布，
# z值在-1.65 ~ 1.65之间时，呈现随机分布。（注：这里说的1.65是临界值，若z值在-1.65 ~ 1.65之间，说明不存在空间自相关性）

# z-socre(standard deviation),P-value(Probability),Confidence level
#<-1.65 or >+1.65.  p<0.10   90%
#<-1.96 or >+1.96.  p<0.05   95%
#<-2.58 or >+2.58.  p<0.01   99%

# The null hypothesis is always there is complete spatial randomness.
```
```{r}
# use the localmoran function to generate I for each ward in the city
I_LWard_Local_count <- points_sf_joined_2 %>%
  pull(plaquecount)%>%
  as.vector()%>%
  localmoran(.,Lward.list)%>%
  as_tibble()

I_LWard_Local_Density <- points_sf_joined_2 %>%
  pull(density)%>%
  as.vector()%>%
  localmoran(.,Lward.list)%>%
  as_tibble()
  
# as_tibble() is a function from the tibble package in R that converts an existing object (like a data frame or matrix) into a tibble. (tidyverse package)

#what does the output (the localMoran object) look like?
slice_head(I_LWard_Local_Density, n=5)
# The slice_head() function is a part of the dplyr package 
# it's used to select the first few rows of a data frame or tibble.

```
```{r}
# copy some of the columns (the I score (column 1) and the z-score standard deviation (column 4)) back into the LondonWards spatialPolygonsDataframe
points_sf_joined_new <- points_sf_joined_2 %>%
  mutate(plaque_count_I= as.numeric(I_LWard_Local_count$Ii))%>%
  mutate(plaque_count_Iz= as.numeric(I_LWard_Local_count$Z.Ii))%>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii))%>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```
## 7.8.5.1 Mapping
```{r}
# first: set the breaks with Confidence level 
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
```

```{r}
# create a new diverging colour brewer palette and reverse the order using rev() (reverse)
library(RColorBrewer)
MoranColours<- rev(brewer.pal(8, "RdGy"))

# brewer.pal() is a function in R that is part of the RColorBrewer package. This function is used to generate color palettes that are specifically designed for thematic mapping. 
```

```{r}
tm_shape(points_sf_joined_new)+
  tm_polygons("plaque_count_Iz",
              style="fixed",
              breaks=breaks1,
              palette= MoranColours,
              midpoint = NA,
              title = "Local Moran's I, Blue Plaques in London")
# style="fixed": Sets the style to "fixed," indicating that you are specifying fixed breaks for the color palette.
# midpoint=NA: Indicates that there is no specific midpoint for the color scale.

#This map shows some areas in the centre of London that have relatively high scores, indicating areas with lots of blue plaques neighbouring other areas with lots of blue plaques.
```
## 7.8.6 Local Getis Ord G∗i
```{r}
# for hot and cold spots
# z-scores: how many standard deviations a value (our value) is away (above or below) from the mean (of the expected values)

# Local Getis Ord G∗i: the local sum (of the neighbourhood) compared to the sum of all features

# The result of Getis Ord G∗ianalysis is an array of Z-scores, one for each pixel [or polygon]
# High Z-scores indicate more intense clustering of high pixel values, indicating hot spots. 
# Low Z-scores indicate more intense clustering of low values, indicating cold spots.
```
```{r}
# calculate by Local G* 
Gi_LWard_Local_Density <- points_sf_joined_2 %>%
  pull(density) %>%
  as.vector()%>%
  localG(., Lward.list)
# check
head(Gi_LWard_Local_Density)
```
```{r}
# increase the column to points_sf_joined_new
points_sf_joined_new <- points_sf_joined_new %>%
  mutate(density_G = as.numeric(Gi_LWard_Local_Density))
```
```{r}
GIColours<- rev(brewer.pal(8, "RdBu"))
# plot on an interactive map
tm_shape(points_sf_joined_new)+
  tm_polygons("density_G",
              style="fixed",
              breaks = breaks1,
              palette= GIColours,
              midpoint= NA,
              title = "Gi*, Blue Plaques in London")
```
# 7.9 Other variables
```{r}
# look at Average GSCE scores
#use head to see what other variables are in the data file

slice_head(points_sf_joined, n=2)
```
```{r}
# Or print out the class of each column like we did in week 2, although we need to drop the geometry.
Datatypelist <- LondonWardsMerged %>% 
  st_drop_geometry()%>%
  summarise_all(class)%>%
  pivot_longer(everything(),
               names_to = "All_variables",
               values_to = "Variable_class")
Datatypelist

````

```{r}
# Local moran'I GSSE
I_LWard_Local_GCSE <- LondonWardsMerged %>%
  arrange(GSS_CODE)%>% # arrange(GSS_CODE) sorts the rows of the data frame based on the values in the GSS_CODE column in ascending order.
  pull(average_gcse_capped_point_scores_2014) %>%
  as.vector()%>%
  localmoran(.,Lward.list)%>%
  as_tibble()

points_sf_joined <- points_sf_joined %>%
  arrange(gss_code)%>%
  mutate(GCSE_LocIz = as.numeric(I_LWard_Local_GCSE$Z.Ii))

tm_shape(points_sf_joined) +
    tm_polygons("GCSE_LocIz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, GCSE Scores")
```

```{r}
# Gi* GSSE
G_LWard_Local_GCSE <- LondonWardsMerged %>%
  dplyr::arrange(GSS_CODE)%>%
  dplyr::pull(average_gcse_capped_point_scores_2014) %>%
  as.vector()%>%
  localmoran(.,Lward.list)%>%
  as_tibble()

points_sf_joined <- points_sf_joined %>%
  dplyr::arrange(gss_code)%>%
  dplyr::mutate(GCSE_LocGiz = as.numeric(G_LWard_Local_GCSE))

tm_shape(points_sf_joined) +
    tm_polygons("GCSE_LocGiz",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Gi*, GCSE Scores")